Kung Faux Pandas \\ \large Simplifying privacy protection

### 2018-07-26 paper/amia.tex
Auto Update from local Sharelatex

\textit{There are many barriers to data access and data sharing, especially in the domain of machine learning on health care data. Legal constraints, such as HIPAA, protect patient privacy but slow access to data and limit reproducibility. We provide a description of an end-to-end system called \emph{Kung Faux Pandas} for easily generating de-identified or synthetic data which is statistically similar to real data but lacks sensitive information. This system focuses on data synthesis and de-identification narrowed to a specific research question to allow for self-service data access without the complexities required to generate an entire population of data that is not needed for a given research project. Kung Faux Pandas is an open source publicly available\footnote{\url{https://github.com/CUD2V/kungfauxpandas}} system that lowers barriers to HIPAA- and GDPR-compliant data sharing for enabling reproducibility and other purposes.}

@@

Independent reproduction and replication of results are critical components of scientific inquiry. Barriers to data access and sharing are the most important impediments to reproducibility in machine learning (ML) research. Health care research has unique challenges because of necessary personal data privacy protections.

Several methods exist to de-identify (remove key identifiers, group individuals, and related anonymization techniques \cite{hippapro}) and synthesize data (create data through algorithmic means, population level statistics, randomization, ideally with preservation of multivariable relationships between records \cite{walonoski_synthea_2018, patki_synthetic_2016, choi_generating_2017}). However, no pipeline currently exists with which ML investigators can routinely \emph{anonymize} or \emph{synthesize} data to facilitate access and sharing. This work addresses that gap.

@@

In the context of ML, sufficient detail is required such that an independent investigator could create the same hardware and software configuration, data used for all experiments, source code, documentation on data and how to run/configure software, and tests that verify the software runs correctly. Once a result can be reproduced, new researchers can then build upon the methods, gather new data for testing/validation, or discover alternative methods to replicate a result.

Many domains in which ML investigators work have legal barriers that make it difficult to enable independent verification of ML models which frequently require enormous amounts of training data. As an example, in the health care domain in the United States, a key law governing health data is the Health Insurance Portability and Accountability Act (HIPAA). HIPAA is backed by significant civil penalties as well as specifics about data security and what can and cannot be done with protected health information \cite{hippaviol}. While there are other domain specific data privacy laws, there are also some broad rules that affect entire countries such as the European Union's General Data Protection Regulation (GDPR). The GDPR puts in place new rules about data ownership and control in an attempt to give individuals more control over their information.

@@

Ideally a ML investigator should develop, test, and validate models using all available data. This is most easily accomplished when they have unfettered access to all data in the problem domain they are working in. Figure~\ref{fig:architecture} shows the novel architecture of KFP: data access through an intermediary that can synthesize data on-demand.

@@

KFP provides multiple methods by which a ML developer can access on-demand faux-data. As shown in Figure~\ref{fig:ui}, a single page web application (SPA) allows for users to directly enter SQL queries, submit them to the connected database, and receive faux-data according to the selected generation method. Data can either be viewed directly in browser or downloaded in csv format for use in a ML process. Alternatively, a Hypertext Transfer Protocol (HTTP) Representational State Transfer (REST) service \cite{w3c_working_group_webservices}, utilized by the SPA, allows for cross language compatible requesting and receiving of faux-data, again via SQL query. In order to facilitate querying and understanding of the data model, the database metadata is presented to the user in a collapsible section. Lastly, for Python based software or languages with an interface to Python, the kungfauxpandas.py file and associated classes can be imported and called natively.

@@

KFP provides a unique contribution to the space of reproducible ML. Through providing ad-hoc faux-data, ML researchers can have lower barriers to accessing private data as well as reduced barriers to sharing faux-data derived from private data. Although one clear area for further research would be to develop plugins for additional faux-data methods, there are several areas that go beyond this simple step that could be researched further: improved code sharing of KFP, addressing institutional specific privacy issues, and evaluation of this system in a ML workflow.

While we have provided software code and a description of our methods, we believe that ML reproducibility should go beyond just making source code available. Peng recommends making code available in any form as an excellent first step towards reproducible research; the next step is to make code available "in a durable non-proprietary format" \cite{peng_reproducible_2011}. This is a level of code sharing rarely achieved in ML research. Currently there is no standard long term format that works for all computational environments. For the Python ecosystem however, the standard for sharing code is via a package on Python Package Index (PyPI). We intend to continue this research to make KFP available as a Python package.

@@

The last area recommended for further research is to deploy and observe the use of KFP in an active ML environment where researchers are working with sensitive datasets. A key limitation that may be identified is that of performance characteristics, particularly in a multi-user setting against a large relational database store. Additionally, as no private data was distributed via this system, actual review by security and compliance groups or institutional review boards may uncover additional requirements to reduce risk of data exposure that could require significant rework or further study.
% include some of this? 
% Lastly, the Synthetic dataset Generation Framework (SGF) by Bindschaedler et al. \cite{Bindschaedler2017} "generates differentially private synthetic data" using a novel criterion called plausible deniability that is applied to a dataset before it is deemed safe to release.  In the case of SGF, several models are built with both real and synthetic data and compared against a gold standard. Their author's analysis showed that the models from synthetic data are close to the same accuracy as models built with real data.

@@

Finally, instead of taking the computational and labor intensive step of full EHR de-identification or synthesis, KFP takes the smaller and more easily implemented step of just-in-time data de-identification or synthesis. As shown in our experiments, if the proper method is selected, data can be quickly generated based on real data that the requester need not have direct access to. We surmise that this tool will improve data exploration and reduce dependency on complex processes required for data access. The result of this system will be improved reproducibility in ML.

>>>>

\textit{There are many barriers to data access and data sharing, especially in the domain of computational research on health care data. Legal constraints, such as HIPAA, protect patient privacy but slow access to data and limit reproducibility. We provide a description of an end-to-end system called \emph{Kung Faux Pandas} for easily generating de-identified or synthetic data which is statistically similar to real data but lacks sensitive information. This system focuses on data synthesis and de-identification narrowed to a specific research question to allow for self-service data access without the complexities required to generate an entire population of data that is not needed for a given research project. Kung Faux Pandas is an open source publicly available\footnote{\url{https://github.com/CUD2V/kungfauxpandas}} system that lowers barriers to HIPAA- and GDPR-compliant data sharing for enabling reproducibility and other purposes.}

@@

Independent reproduction and replication of results are critical components of scientific inquiry. Barriers to data access and sharing are the most important impediments to reproducibility in health data science research. Health care research has unique challenges because of necessary personal data privacy protections.

Several methods exist to de-identify (remove key identifiers, group individuals, and related anonymization techniques \cite{hippapro}) and synthesize data (create data through algorithmic means, population level statistics, randomization, ideally with preservation of multivariable relationships between records \cite{walonoski_synthea_2018, patki_synthetic_2016, choi_generating_2017}). However, no pipeline currently exists with which computational investigators can routinely \emph{anonymize} or \emph{synthesize} data to facilitate access and sharing. This work addresses that gap.

@@

In the context of machine learning (ML), sufficient detail is required such that an independent investigator could create the same hardware and software configuration, data used for all experiments, source code, documentation on data and how to run/configure software, and tests that verify the software runs correctly. Once a result can be reproduced, new researchers can then build upon the methods, gather new data for testing/validation, or discover alternative methods to replicate a result.

Many domains in which health data science investigators work have legal barriers that make it difficult to enable independent verification of ML models which frequently require enormous amounts of training data. As an example, in the health care domain in the United States, a key law governing health data is the Health Insurance Portability and Accountability Act (HIPAA). HIPAA is backed by significant civil penalties as well as specifics about data security and what can and cannot be done with protected health information \cite{hippaviol}. While there are other domain specific data privacy laws, there are also some broad rules that affect entire countries such as the European Union's General Data Protection Regulation (GDPR). The GDPR puts in place new rules about data ownership and control in an attempt to give individuals more control over their information.

@@

Ideally a computational investigator should develop, test, and validate models using all available data. This is most easily accomplished when they have unfettered access to all data in the problem domain they are working in. Figure~\ref{fig:architecture} shows the novel architecture of KFP: data access through an intermediary that can synthesize data on-demand.

@@

KFP provides multiple methods by which a health data scientist can access on-demand faux-data. As shown in Figure~\ref{fig:ui}, a single page web application (SPA) allows for users to directly enter SQL queries, submit them to the connected database, and receive faux-data according to the selected generation method. Data can either be viewed directly in browser or downloaded in csv format for use in a ML process. Alternatively, a Hypertext Transfer Protocol (HTTP) Representational State Transfer (REST) service \cite{w3c_working_group_webservices}, utilized by the SPA, allows for cross language compatible requesting and receiving of faux-data, again via SQL query. In order to facilitate querying and understanding of the data model, the database metadata is presented to the user in a collapsible section. Lastly, for Python based software or languages with an interface to Python, the kungfauxpandas.py file and associated classes can be imported and called natively.

@@

KFP provides a unique contribution to the space of reproducible computational research. Through providing ad-hoc faux-data, health data scientists can have lower barriers to accessing private data as well as reduced barriers to sharing faux-data derived from private data. Although one clear area for further research would be to develop plugins for additional faux-data methods, there are several areas that go beyond this simple step that could be researched further: improved code sharing of KFP, addressing institutional specific privacy issues, and evaluation of this system in a ML workflow.

While we have provided software code and a description of our methods, we believe that computational reproducibility should go beyond just making source code available. Peng recommends making code available in any form as an excellent first step towards reproducible research; the next step is to make code available "in a durable non-proprietary format" \cite{peng_reproducible_2011}. This is a level of code sharing rarely achieved in computation based research. Currently there is no standard long term format that works for all computational environments. For the Python ecosystem however, the standard for sharing code is via a package on Python Package Index (PyPI). We intend to continue this research to make KFP available as a Python package.

@@

The last area recommended for further research is to deploy and observe the use of KFP in an active health data science environment where researchers are working with sensitive datasets. A key limitation that may be identified is that of performance characteristics, particularly in a multi-user setting against a large relational database store. Additionally, as no private data was distributed via this system, actual review by security and compliance groups or institutional review boards may uncover additional requirements to reduce risk of data exposure that could require significant rework or further study.
% include some of this? 
% Lastly, the Synthetic dataset Generation Framework (SGF) by Bindschaedler et al. \cite{Bindschaedler2017} "generates differentially private synthetic data" using a novel criterion called plausible deniability that is applied to a dataset before it is deemed safe to release.  In the case of SGF, several models are built with both real and synthetic data and compared against a gold standard. Their author's analysis showed that the models from synthetic data are close to the same accuracy as models built with real data.

@@

Finally, instead of taking the computational and labor intensive step of full EHR de-identification or synthesis, KFP takes the smaller and more easily implemented step of just-in-time data de-identification or synthesis. As shown in our experiments, if the proper method is selected, data can be quickly generated based on real data that the requester need not have direct access to. We surmise that this tool will improve data exploration and reduce dependency on complex processes required for data access. The result of this system will be improved reproducibility in health data science.
